Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2015-09-22 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add example of snappyHexMesh behaviour including diagram

\documentclass[draft]{ametsoc}

@@

There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid created using a shaving method, as described by \citet{adcroft1997}.  First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.  Second, the utility does not create new points necessary for pentagonal cells.

@@

\subsection{Finite difference linear advection model}
The second model is a modified version of the linear advection model first used by \citet{schaer2002}.  It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme:
\begin{align}
	\frac{\partial \phi_i}{\partial x} \approx \frac{1}{\Delta x} \left( \frac{1}{12} \phi_{i-2} - \frac{2}{3} \phi_{i-1} + \frac{2}{3} \phi_{i+1} - \frac{1}{12} \phi_{i+2} \right)
\end{align}

@@

The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.  The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.  This sensitivity can be explained by realising that increased non-orthogonality or flow misalignment tends to increase errors.  To measure non-orthogonality, consider the total variation, $TV$, of the discrete terrain profile:
\begin{align}
	TV = \sum_i | h(x_{i+1}) - h(x_i) |
\end{align}
where $x_i$ is the horizontal location of a vertex on the lower boundary.  For a domain width of \SI{301}{\kilo\meter}, $TV = \SI{18095}{\meter}$, but for a width of \SI{300}{\kilo\meter}, $TV = \SI{18964}{\meter}$.
The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.

@@

>>>>

\documentclass[twocol]{ametsoc}
\journal{mwr}

@@

There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid created using a shaving method, as described by \citet{adcroft1997}.  First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.  Second, it does not create new points necessary for pentagonal cells.  An example of these effects are shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.  The resultant mesh, shown by thick dashed lines, contains triangular and quadrangular cells.  With a typical shaving method, cell $c$ would be pentagonal, as shown by the dotted line.  However, \texttt{snappyHexMesh} creates a quadrilateral cell.

@@

\subsection{Finite difference linear advection model}
The second model is a modified version of the linear advection model first used by \citet{schaer2002}.  It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme:
\begin{align}
	\frac{\partial \phi_i}{\partial x} \approx \frac{1}{\Delta x} \left( \frac{1}{12} \phi_{i-2} - \frac{2}{3} \phi_{i-1} + \frac{2}{3} \phi_{i+1} - \frac{1}{12} \phi_{i+2} \right)
\end{align}
and similarly for \(\partial \phi_i / \partial z\).

@@

The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.  The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.  The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.  It is likely that changes in the domain width affect the wave power spectrum of the discrete terrain profile.  High-frequency oscillations persist in the centred linear scheme but are damped by the upwind-biased scheme.  This is an area for further study.

@@

\begin{figure*}
	\centering
	\includegraphics{../fig-snappyHexMesh/fig-snappyHexMesh.pdf}
	%
	\caption{Example behaviour of the \texttt{snappyHexMesh} tool in cut cell grid creation.  A terrain surface denoted by a thick grey line is intersected with an orthogonal grid shown by thin solid lines.  Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.  The resultant cut cell grid is shown by thick dashed lines.  With a typical cut cell technique, cell $c$ would be a pentagon, as shown by the dotted line, but \texttt{snappyHexMesh} produces a quadrangular cell.}
	\label{fig:snappyHexMesh}
\end{figure*}
